---
title: "Project EDA-Red Wines"
output: html_document
---

```{r setup, include=FALSE}
# knitr: Suppress code/messages/warnings 
#        Set default plot options and center them
knitr::opts_chunk$set(fig.width=9,fig.height=5,fig.path='Figs/',
                      fig.align='center',tidy=TRUE,
                      echo=FALSE,warning=FALSE,message=FALSE)
```

```{r cars}
summary(cars)
```

## Including Plots

Red Wine Quality by ZAID SAAD ALZUAIBER
========================================================
  
#introduction:
  This tidy data set contains 1,599 red wines with 11 variables on the chemical 
properties of the wine. At least 3 wine experts rated the quality of each wine, 
providing a rating between 0 (very bad) and 10 (very excellent).



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

library(gridExtra) 
library(ggplot2)
library(knitr)
library(dplyr)

```

```{r echo=FALSE, Load_the_Data}

pf <- read.csv('wineQualityReds.csv')

```

```{r}

# Univariate Plots Section

str(pf)
```
```{r}
names(pf)
```
```{r}
summary(pf)
```


```{r ,quality}

# bar Quality

ggplot(aes(x=quality), data = pf)+
  geom_bar( fill = "#993366") +
  xlab("Quality") +
  ggtitle("Quality")


summary(pf$quality)
```

We noticed the red win (qualty) is normaly distrbuted and mean around 5.0


```{r ,alcohol}
#bar chart
ggplot(aes(x=alcohol), data = pf)+
  geom_bar( fill = "#993366") +
  xlab("alcohol % ") +
    ggtitle("alcohol %")


summary(pf$alcohol)
```

We noticed the red win (alcohol) is a right skewed and mean  around 10


```{r ,total.sulfur.dioxide}
#total.sulfur.dioxide

ggplot(aes(x=total.sulfur.dioxide), data = pf)+
  geom_histogram( bins = 50,fill = "#993366",color="white") +
  xlab("total.sulfur.dioxide") +
      ggtitle("total.sulfur.dioxide")


summary(pf$total.sulfur.dioxide)
```

We noticed the red win (total.sulfur.dioxide) is a right skewed and mean around 46


```{r }

#fixed.acidity

ggplot(aes(x=fixed.acidity), data = pf)+
  geom_bar( fill = "#993366") +
  xlab("fixed.acidity") +
      ggtitle("total.sulfur.dioxide")

summary(pf$fixed.acidity)
```

We notice the red win (fixed.acidity) is a right skewed and mean around 8 


```{r }
#volatile.acidity

ggplot(aes(x=volatile.acidity), data = pf)+
  geom_histogram( color="white",fill = "#993366") +
  xlab("volatile.acidity") +
      ggtitle("volatile.acidity")

summary(pf$volatile.acidity)
```

We noticed the red win (volatile.acidity) is a right skewed and mean around 0.25 and
there is a few outlier


```{r }
#citric.acid

ggplot(aes(x=citric.acid), data = pf)+
  geom_bar( fill = "#993366") +
  xlab("citric.acid") +
      ggtitle("citric.acid")

summary(pf$citric.acid)
```

We notice the red win (citric.acid) there is one oulier is "1" and mean around 0.27


```{r }
#residual.sugar

ggplot(aes(x=residual.sugar), data = pf)+
  geom_histogram(bin=50,color="white", fill = "#993366") +
  xlab("residual.sugar") +
      ggtitle("citric.acid")

summary(pf$residual.sugar)
```

We notice the red win (residual.sugar) is a right skewed  and mean  around 2.53 


```{r }
#chlorides

ggplot(aes(x=chlorides), data = pf)+
  geom_histogram(color="white", fill = "#993366") +
  xlab("chlorides") +
      ggtitle("chlorides")

summary(pf$chlorides)
```

We notice the red win (chlorides) there is many outliers  and mean  around 0.08

```{r }
#density

ggplot(aes(x=density), data = pf)+
  geom_histogram( color="white",fill = "#993366") +
  xlab("density") +
      ggtitle("density")

summary(pf$density)
```
We notice the red win (density) normal distrubtions  and mean  around 0.99


```{r }
#sulphates

ggplot(aes(x=sulphates), data = pf)+
  geom_bar( fill = "#993366") +
  xlab("sulphates") +
      ggtitle("sulphates")

summary(pf$sulphates)
```
We notice the red win (sulphates) is a right skewed   and mean  around 0.65


```{r }

names(pf)
```

# Univariate Analysis


### What is the structure of your dataset?
there are 1599 wines and 13 variables  ("X","fixed.acidity","volatile.acidity","citric.acid"         
"residual.sugar","chlorides","free.sulfur.dioxide","total.sulfur.dioxide"
"density","pH","sulphates","alcohol","quality" )
### What is/are the main feature(s) of interest in your dataset?
The main feature is "quality",because it will help me to decide which product can we predict it's better
and we campare with other factors
### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
i think  density and alcohol 
### Did you create any new variables from existing variables in the dataset?
No
### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form 
of the data? If so, why did you do this?
No I didin't' the data are tidy

### Bivariate Plots Section

```{r}
#total.sulfur.dioxide and free.sulfur.dioxide


ggplot(aes(x= total.sulfur.dioxide, y = free.sulfur.dioxide) , data = pf) +
  geom_jitter(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1) +
  scale_color_brewer(type='seq')+
    ggtitle("total.sulfur.dioxide and free.sulfur.dioxide")


```
As we see after we compared between total.sulfur.dioxide and free.sulfur.dioxide we can see the free.sulfur.dioxide with wines higher when it's close 0


```{r}
#alcohol and quality

ggplot(aes(x= alcohol, y = quality) , data = pf) +
  geom_point(alpha =0.5, size = 4) +
  geom_smooth(method = "lm", se = FALSE,size=1) +
  xlab("alcohol %")+
  scale_color_brewer(type='seq')+
    ggtitle("alcohol and quality")

```
as we can see once the alcohol getting high percentage it will be good quality 

```{r}
#quality and density


ggplot(aes(x= quality, y = density) , data = pf) +
  geom_rug(alpha = 0.3, size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_brewer(type='seq')+
  xlab(density~(g/cm^3))
    ggtitle("quality and density")

```
As we can see the density it's not making significant impact on quality
```{r}
#pH and fixed.acidity

ggplot(aes(x= pH, y = fixed.acidity) , data = pf) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1) +
  scale_color_brewer(type='seq')+
    ggtitle("quality and density")

```

As four graphs above it shows the correletion between 2 variables which we compared.
I choosed the chemical features that might be has a significant correlation with wine quality.



# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
as the graphs above we saw many correletion between varieables like ( fixed.acidity and ph ) around 3,3

  
### Did you observe any interesting relationships between the other features \
  (not the main feature(s) of interest)?
YES, we found relation between( alcohol and quality) the quality will be good once the quantity of
alcohol has more
### What was the strongest relationship you found?
alcohol and quality as we mentioned earlier.  
  
  
  
# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
#Alcohol and citric.acid, by Quality

ggplot(aes(x = alcohol, y = citric.acid, color = factor(quality)), data = pf) +
  geom_point(alpha = 0.8, size = 5) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  theme_dark() +
  xlab("alcohol%")+
  scale_color_brewer(type='seq')+
  ggtitle("Alcohol and citric.acid, by Quality")



```
As we can see there is relation between quality and alcohol and citric.acid 
once alcohol increased it shows the quality increased

```{r}

#Alcohol and citric.acid, by Quality

ggplot(pf, aes(x = sulphates, y = alcohol, colour=factor(quality))) + 
  geom_point(alpha = 0.8, size = 4) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  theme_dark() +
  ylab("alcohol%")+
  scale_color_brewer(type='seq')+
  ggtitle("sulphates and alcohol, by Quality")
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
as we mentioned in graph above when alcohol got higher the quality will increased

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.
No



# Final Plots and Summary


### Plot One

```{r}
#quality

ggplot(aes(x=quality), data = pf)+
  geom_bar( fill = "#993366") +
  xlab("quality") +
  ggtitle("Quality")

```
### Description One
The graph shows the quality of wine and we can see almost wines have quality around 5


### Plot Two
```{r echo=FALSE, Plot_Two}
#alcohol and quality

ggplot(aes(x= alcohol, y = quality) , data = pf) +
  geom_smooth(alpha = 0.3, size = 3) +
  xlab("alcohol%")+
  ggtitle("alcohol and quality")
```
### Description Two
As we see the correlation between( alcohol and quality) are positve



### Plot Three
```{r, Plot_Three}
#pH and alcohol, by Quality

ggplot(aes(x = pH, y = alcohol, color = factor(quality)), data = pf) +
  geom_jitter(alpha = 0.5, size = 4) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  theme_dark() +
  ylab("alcohol%")+
  scale_color_brewer(type='seq')+
  ggtitle("pH and alcohol, by Quality")

```

### Description Three
High quality wines when higher alcohol and high pH




------
  
# Reflection
  
The red wines dataset contains 1599 observations with 13 variables.
After analyzing and exploring data such as what I expected the alcohol 
percentage correlated to quality, but I was expect the quality will be 
good in the higher acidity.
I think we need more features to make sure what impacts on wines like (fermentation in terms of the time and the process)
it's diffculty to learn new langyage but i found it easier than python in plot and helpful 



